Synthesizing Symmetric Lenses

### 2018-06-29 papers/SymmetricLenses/main.tex
minor wording changes

Lens languages, like any domain-specific language, also cut down the space of
possible programs one can write, and the smaller the space of programs,
the easier it is to find the program one is looking for.  
%This feature can be a negative, if the desired transformation
%falls outside of the scope provided by the language, but can also be positive, as complex
%properties (like invertability) can be enforced automatically.
%Another advantage of reducing the
%set of programs is that doing so can facilitate program synthesis, which is essentially
%a search through the space of possible computations that can be expressed as a lens.
Recent research~\cite{optician}
has exploited this fact by demonstrating it is possible to synthesize relatively
large, complex lenses from specifications.
In particular, given types that describe source and target formats, there are synthesis
algorithms that will find bijective lenses~\cite{optician}
for medium-sized files, such as those housing linux
application configurations.  \ksf{why medium sized?  You can have very
large files with simple format descriptions that are bijective and for
which our previous systems would have worked just fine.} Oftentimes, these algorithms find such lenses in less than
a second, and without the aid of any examples at all, with just types describing the source and
target formats.  \dpw{I would cite the ICFP paper (cite string: maina+:quotient-synthesis) as to appear but I'm not sure how to do that
without possibly violating the spirit of double-blind reviewing. ie, if it hasn't appeared yet,
how could we know the content if we aren't the authors?  Mentioning it in the 3rd person does not
seem to suffice ...} \ksf{The ICFP 2018 website lists the paper; I
think it is okay to cite it.  The reviewer can assume we contacted the
authors for a copy.  We could ask Stephanie for a ruling if we want to
be sure.  One of the principles of double blind as adopted by SIGPLAN
is not to reduce the quality of the paper; omitting this citation
woudl have that affect.}

@@

Initially, we hypothesized that extending the algorithms
of Miltner \ETAL~\cite{miltner+:bijective-synthesis} to operate over symmetric lenses
as opposed to bijective lenses 
would be relatively straightforward:  Simply replace one set of combinators with another
and execute a search with similar heuristics.  However, in our early,
naive attempts to do so, we ran
into two different kinds of difficulties. \dpw{I got into the weeds a little bit there.
The goal was to set up the idea that there are ``two parts'' to a symmetric
lens transformation: the information from $s$ and the information from $t$.  In
a future paragraph in the intro we are going to explain that the central heuristic
involves minimizing the information from $t$.}

@@

In order to characterize the expressive power of simple symmetric lenses
theoretically, we prove simple symmetric lenses
include all (and only) those symmetric lenses that satisfy
an intuitive property called \emph{forgetfulness}.  In other words, they are the
complete set of symmetric lenses that do not require a ``memory'' to retain data
from past instances of $S$ or $T$ when making a round trip.\dpw{fix that to make the intuitive
explaination sufficiently accurate}
From a pragmatic perspective, we justify our design choices by extending
the Boomerang implementation with symmetric lenses (they had never before been implemented),
applying them successfully to a range of real-world applications,
and demonstrating that they are compatible, not just with basic bijective combinators,
but with a rich set of advanced features including quotient lenses and matching lenses.

>>>>

Lens languages, like any domain-specific language, also cut down the space of
possible programs one can write, and the smaller the space of programs,
the easier it is to find the program one is looking for.  
%This feature can be a negative, if the desired transformation
%falls outside of the scope provided by the language, but can also be positive, as complex
%properties (like invertability) can be enforced automatically.
%Another advantage of reducing the
%set of programs is that doing so can facilitate program synthesis, which is essentially
%a search through the space of possible computations that can be expressed as a lens.
Recent research~\cite{optician}
has exploited this fact by demonstrating that it is possible to synthesize relatively
large, complex lenses from specifications.
In particular, given types that describe source and target formats, there are synthesis
algorithms that will find bijective lenses~\cite{optician}
for medium-sized files, such as those housing linux
application configurations.  \ksf{why medium sized?  You can have very
large files with simple format descriptions that are bijective and for
which our previous systems would have worked just fine.} Oftentimes, these algorithms find such lenses in less than
a second, and without the aid of any examples at all, with just types describing the source and
target formats.  \dpw{I would cite the ICFP paper (cite string: maina+:quotient-synthesis) as to appear but I'm not sure how to do that
without possibly violating the spirit of double-blind reviewing. ie, if it hasn't appeared yet,
how could we know the content if we aren't the authors?  Mentioning it in the 3rd person does not
seem to suffice ...} \ksf{The ICFP 2018 website lists the paper; I
think it is okay to cite it.  The reviewer can assume we contacted the
authors for a copy.  We could ask Stephanie for a ruling if we want to
be sure.  One of the principles of double blind as adopted by SIGPLAN
is not to reduce the quality of the paper; omitting this citation
woudl have that affect.}\saz{Stephanie says we should cite it in the third
person, even if that's a bit awkward, but otherwise citing it is OK.}

@@

Initially, we hypothesized that extending the synthesis algorithms
of Miltner \ETAL~\cite{miltner+:bijective-synthesis} to operate over symmetric lenses
as opposed to bijective lenses 
would be relatively straightforward:  Simply replace one set of combinators with another
and execute a search with similar heuristics.  However, in our early,
naive attempts to do so, we ran
into two different kinds of difficulties. \dpw{I got into the weeds a little bit there.
The goal was to set up the idea that there are ``two parts'' to a symmetric
lens transformation: the information from $s$ and the information from $t$.  In
a future paragraph in the intro we are going to explain that the central heuristic
involves minimizing the information from $t$.}

@@

In order to characterize the expressive power of simple symmetric lenses
theoretically, we prove that simple symmetric lenses
include all (and only) those symmetric lenses that satisfy
an intuitive property called \emph{forgetfulness}.  In other words, they are the
complete set of symmetric lenses that do not require a ``memory'' to retain data
from past instances of $S$ or $T$ when making a round trip.\dpw{fix that to make the intuitive
explaination sufficiently accurate}
From a pragmatic perspective, we justify our design choices by extending
the Boomerang implementation with symmetric lenses (they had never before been implemented),
applying them successfully to a range of real-world applications,
and demonstrating that they are compatible, not just with basic bijective combinators,
but with a rich set of advanced features including quotient lenses and matching lenses.
